---
title: "GISPR Module 9 — Statistical Modeling, reticulate, and Spatial Web Services"
subtitle: "Course: GIS Spatial Analysis with Python and R (GISPR)"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
engine: knitr
---

<!--
GISPR Module 9 lab — built as a Quarto (.qmd) document rather than the .ipynb
format used in Modules 1–8. That's a deliberate change, not an oversight: the
whole point of this module is one working session where R and Python hand
data back and forth without switching kernels, and Quarto's knitr engine does
exactly that automatically wherever `reticulate` is installed — an R chunk's
objects show up in the next Python chunk as `r.<name>`, and a Python chunk's
objects show up in the next R chunk as `py$<name>`. `engine: knitr` in the
YAML above locks that behavior in (Quarto would otherwise default a
Python-containing .qmd to a separate Jupyter engine, which is the
kernel-switching model this module exists to replace).

Open this file in Positron, click Render (or run chunks interactively with
Cmd/Ctrl+Enter) — there's no kernel picker to think about. Positron runs the
whole document as one R session with Python embedded inside it via
reticulate.
-->

# What you'll accomplish this module

By the end of this notebook you'll be able to:

| # | Learning outcome | Where |
|---|---|---|
| 1 | Design an integrated workflow: apply statistical modeling / machine learning libraries (`scikit-learn`, `tidymodels`) in R and Python to a real spatial prediction problem, end to end | Sections 3–4 |
| 2 | Work across R and Python: transfer data and execute functions between R and Python within a single workflow using `reticulate` and Positron | Sections 1, 3, 5 |
| 3 | Pull live spatial data: access and integrate spatial data from web services — WMS/WFS, REST APIs, and IoT data streams — into an R/Python spatial workflow | Section 2 |

### The framing problem for tonight

> **Where should the city add its next 10 low-cost air-quality sensors?**
> The city has a sparse network of PM2.5 monitors — good enough for regulatory
> compliance, not good enough to say whether the neighborhood next to the
> rail yard is breathing worse air than the one next to the park. You have
> live sensor readings from a REST API, land-cover / traffic / elevation
> covariates from a WFS, and a budget for 10 more sensors. Build a model that
> predicts PM2.5 everywhere from what a sensor *can* see, then rank the
> unmonitored locations most likely to be under-measuring real exposure.

Nothing here is new spatial theory — you've done spatial joins (Module 4),
raster/vector integration (Modules 5–6, 8), and mapping (Module 7). Module 9
adds three new pieces to the pipeline: a *statistical model* standing in for
the "integrate" stage, a *language bridge* so you're not locked into whichever
library happens to be in R or Python, and *live web services* replacing the
downloaded file as your data source.

### Before you start — confirm your R ↔ Python bridge

Run this chunk first. If it errors, fix it now — don't wait for the Guided Lab.

```{r}
#| label: env-check

library(reticulate)
py_config()
# Expect: a Python path + version printed above.

py_run_string("print('Python is connected')")
# Expect: the printed confirmation string, with no error in between.
```

**If it failed:** the most common cause is that `reticulate` auto-created a
bare `r-reticulate` conda environment with no GIS/ML packages installed.
Two fixes:

```{r}
#| eval: false

# Fix 1 — let reticulate build a working environment for you
reticulate::install_miniconda()

# Fix 2 — point reticulate at Python you already trust (e.g. ArcGIS Pro's,
# or a conda env where you've already installed pandas/scikit-learn)
reticulate::use_python("path/to/arcgispro-py3/python.exe")
```

Re-run the `env-check` chunk above after either fix, and confirm it prints
cleanly before moving on.

---

# Section 1 — The Bridge: How `reticulate` Actually Works {#sec-bridge}

*(Concept Input, ~10 min of tonight's live-coding segment)*

`reticulate` embeds a Python interpreter inside your R session. You aren't
switching between two programs — one process, two languages, objects passed
back and forth by name.

## 1a — Four ways to use Python from R

| Method | What it does | When you'd use it |
|---|---|---|
| `source_python("script.py")` | Runs a `.py` file, pulls its top-level functions/variables into R | You already have a Python module and just want to call its functions |
| `py_run_string("...")` | Runs a string of Python code in the shared session | A quick one-off Python snippet inside an R script |
| `repl_python()` | Drops you into an interactive Python prompt inside your R console; `exit` returns to R | Exploring interactively before you commit code to a chunk |
| Quarto `{python}` chunks (what this document uses) | Each chunk is Python, but shares the reticulate session with the R chunks around it | Writing a document where both languages appear as first-class code, not a side quest |

## 1b — Type conversion, the short version

| R | Python | Notes |
|---|---|---|
| `data.frame` | `pandas.DataFrame` | Converts automatically, column types mostly preserved |
| `list` (named) | `dict` | Named list ↔ dict is the default two-way mapping |
| `list` (unnamed) | `list` | |
| numeric vector | NumPy array / list | Length-1 vectors become Python scalars |
| `NULL` | `None` | |
| `sf` object | *(no automatic conversion)* | Convert geometry to WKT or GeoJSON text first — reticulate doesn't know GeoPandas' GeoDataFrame; you're the bridge for spatial types (see Section 3) |

## 1c — Live coding: pass an object each direction

```{r}
#| label: bridge-demo-r1

library(reticulate)

# An R data.frame — nothing python-flavored about it yet
sensor_counts_r <- data.frame(
  land_use = c("industrial", "residential", "commercial", "park"),
  n_sensors = c(2L, 5L, 3L, 1L)
)
sensor_counts_r
```

```{python}
#| label: bridge-demo-py

# `r.sensor_counts_r` reaches back into the R chunk above — the data.frame
# arrives already converted to a pandas DataFrame, no export/import step.
import pandas as pd

sensor_counts_py = r.sensor_counts_r
sensor_counts_py["sensors_per_10k_pop"] = sensor_counts_py["n_sensors"] * 1.7  # toy scaling
sensor_counts_py
```

```{r}
#| label: bridge-demo-r2

# `py$sensor_counts_py` reaches forward into the Python chunk above —
# the pandas DataFrame arrives back as an R data.frame, edits and all.
sensor_counts_back <- py$sensor_counts_py
str(sensor_counts_back)
```

That round trip — R → Python → R, no files written, no kernel switched — is
the entire objective in miniature. Everything else this module does is that
same handoff applied to spatial data and a statistical model.

## 1d — When to reach for the bridge (and when not to)

| Reach for `reticulate` when… | Just pick one language when… |
|---|---|
| The library you need only exists in the other language (e.g. a specific `scikit-learn` estimator, or an R spatial-statistics package with no Python equivalent) | Both languages have a perfectly good option and switching buys you nothing |
| You want to compare two implementations of the same idea (tidymodels vs. scikit-learn) side by side in one document | The task is small enough that the bridge's overhead isn't worth explaining to whoever reads this next |
| You're prototyping in Python but the rest of your team's pipeline is R (or vice versa) | You're handing this off to someone who only knows one of the two languages — don't make them install both toolchains to read your code |

---

# Section 2 — Pulling Data From Web Services {#sec-webservices}

*(Concept Input, remainder of the live-coding segment)*

Three request/response patterns show up under the umbrella of "web services,"
and Module 9's objective names all three:

| Pattern | What you get back | Example this lab uses |
|---|---|---|
| **WMS** (Web Map Service) | A rendered image — a picture, not features you can query or join | *(not used for modeling — mentioned for contrast; you'd reach for this only to display a reference basemap layer)* |
| **WFS** (Web Feature Service) / generic **REST API** | Actual features — geometries + attributes you can filter, join, and feed to a model | Land-cover, traffic-distance, and elevation covariates |
| **IoT / SensorThings-style stream** | Time-stamped observations from physical devices, usually JSON over REST | The PM2.5 sensor network readings |

## 2a — Why this notebook doesn't hit a real endpoint

Real WFS/REST endpoints go down, change schemas, and rate-limit — bad
properties for a notebook forty students run at once. Following the same
"synthetic-but-realistic" convention as Modules 3–8, this lab builds a
synthetic REST-style sensor feed and a synthetic WFS-style covariate layer
that look exactly like what the real services would hand back, so the *shape*
of the code — the part that matters — is identical to hitting the live
services. Swap the base URL back in and the rest of the pipeline runs
unchanged; that swap-in point is called out explicitly below.

```{python}
#| label: synth-sensors

# [Python] Step 0a — a synthetic REST/IoT-style PM2.5 sensor feed
#
# A real call would look like:
#   import requests
#   resp = requests.get(
#       "https://api.citysensors.example.gov/v1/pm25/current",
#       params={"bbox": "-123.30,44.55,-123.20,44.63"},
#   )
#   sensors_json = resp.json()
#
# We build the equivalent payload locally instead — same JSON shape
# (a list of records with lon/lat/value/timestamp) a real IoT REST API
# would return.
import numpy as np
import pandas as pd

rng = np.random.default_rng(9)
n_sensors = 22

lon = rng.uniform(-123.29, -123.21, n_sensors)
lat = rng.uniform(44.56, 44.62, n_sensors)

# PM2.5 (µg/m3): higher near the low-lon/low-lat corner (stand-in for an
# industrial/traffic corridor), lower toward the high-lon/high-lat corner
# (stand-in for a park/residential edge) — a believable exposure gradient.
base = 14 - 60 * (lon + 123.29) - 40 * (lat - 44.56)
pm25 = np.clip(base + rng.normal(0, 1.5, n_sensors), 3, 30)

sensors_json = [
    {
        "sensor_id": f"S-{i:03d}",
        "lon": round(float(lon[i]), 5),
        "lat": round(float(lat[i]), 5),
        "pm25": round(float(pm25[i]), 1),
        "timestamp": "2026-07-15T18:00:00Z",
    }
    for i in range(n_sensors)
]

sensors_df = pd.DataFrame(sensors_json)
print(f"Pulled {len(sensors_df)} sensor readings (synthetic REST/IoT payload)")
sensors_df.head()
```

```{r}
#| label: synth-covariates

# [R] Step 0b — a synthetic WFS-style covariate layer
#
# A real call would look like:
#   library(httr2)
#   resp <- request("https://gis.city.example.gov/geoserver/wfs") |>
#     req_url_query(service = "WFS", version = "2.0.0", request = "GetFeature",
#                   typeNames = "city:landcover_grid", outputFormat = "application/json") |>
#     req_perform()
#   covariates_sf <- sf::read_sf(resp_body_string(resp))
#
# We build the equivalent sf object locally — a regular grid of cells with
# percent impervious surface, distance to the nearest major road (m), and
# elevation (m) — the same three columns a real land-cover/traffic/DEM-derived
# WFS layer would carry, in the same projected CRS the course has used since
# Module 3 (EPSG:32610, UTM Zone 10N).
library(sf)
library(dplyr)

set.seed(9)
ncell <- 18
xmin <- 500000; ymin <- 4937000; cellsize <- 400   # UTM 10N, arbitrary origin near Corvallis-scale extent

grid_geom <- st_make_grid(
  st_bbox(c(xmin = xmin, ymin = ymin,
            xmax = xmin + ncell * cellsize, ymax = ymin + ncell * cellsize)),
  cellsize = cellsize, what = "polygons"
)

covariates_sf <- st_sf(
  cell_id = seq_along(grid_geom),
  geometry = grid_geom,
  crs = 32610
)

centroids <- st_coordinates(st_centroid(covariates_sf))
# Same gradient logic as the Python sensor readings above, so the covariates
# actually explain the exposure surface rather than being unrelated noise.
frac_x <- (centroids[, "X"] - xmin) / (ncell * cellsize)
frac_y <- (centroids[, "Y"] - ymin) / (ncell * cellsize)

covariates_sf <- covariates_sf |>
  mutate(
    pct_impervious = pmax(0, pmin(100, 85 - 70 * frac_x - 50 * frac_y + rnorm(ncell^2, 0, 6))),
    dist_major_road_m = pmax(20, 900 * frac_x + 600 * frac_y + rnorm(ncell^2, 0, 80)),
    elevation_m = 60 + 40 * frac_y + rnorm(ncell^2, 0, 5)
  )

st_write(covariates_sf, "output/covariates.gpkg", delete_dsn = TRUE, quiet = TRUE)
cat("Wrote", nrow(covariates_sf), "covariate grid cells to output/covariates.gpkg\n")
head(st_drop_geometry(covariates_sf))
```

## 2b — SensorThings and the IoT piece of the objective

The sensor feed above is a snapshot — one reading per device. A real IoT
platform (OGC's **SensorThings API** is the standard named in this module's
required reading) exposes each device as a `Thing` with a time series of
`Observations`, so you can ask for "the last 24 hours of PM2.5 from sensor
S-014" rather than just its current value. That's the same REST pattern with
one more nesting level — worth knowing the vocabulary (`Thing`,
`Datastream`, `Observation`) even though this lab works from single readings
to keep the modeling steps focused.

---

# Section 3 — Guided Lab: Build the Air-Quality Exposure Model {#sec-lab}

*(Guided Lab, ~50–60 min)*

**Goal:** a `predict_exposure()` pipeline that takes the sensor readings and
covariate grid from Section 2, fits a spatial model, and ranks the
unmonitored cells most likely to be under-measured — the shortlist for the
city's next 10 sensors.

## Step 1 — Spatial join: attach covariates to each sensor

```{r}
#| label: lab-step1

# [R] Step 1 — join sensor points to the covariate cell they fall in
library(sf)
library(dplyr)

sensors_sf <- st_as_sf(
  py$sensors_df,                     # <- the pandas DataFrame from Section 2, back across the bridge
  coords = c("lon", "lat"), crs = 4326
) |>
  st_transform(32610)                # match the covariate grid's CRS before joining — Module 3's rule

sensors_joined <- st_join(sensors_sf, covariates_sf, join = st_within)

cat("Sensors joined to covariates:", sum(!is.na(sensors_joined$cell_id)),
    "of", nrow(sensors_joined), "\n")
head(st_drop_geometry(sensors_joined))
```

## Step 2 — The naive-split trap (why this isn't just `train_test_split()`)

Before fitting anything: your 22 sensors aren't independent observations.
Two sensors four blocks apart see correlated air — Tobler's First Law again,
same as the concept slide from tonight's warm-up. Split them randomly into
train/test and the test sensors likely sit right next to a training sensor,
so the model "passes" by memorizing its neighbor rather than learning the
land-cover/traffic/elevation relationship. Watch it happen:

```{r}
#| label: lab-step2-naive

# [R] Step 2a — naive random cross-validation (the wrong way, done on purpose)
library(tidymodels)

model_data <- sensors_joined |>
  st_drop_geometry() |>
  filter(!is.na(cell_id)) |>
  select(pm25, pct_impervious, dist_major_road_m, elevation_m)

set.seed(9)
naive_folds <- vfold_cv(model_data, v = 5)

rf_spec <- rand_forest(trees = 500) |>
  set_engine("ranger") |>
  set_mode("regression")

rf_wf <- workflow() |>
  add_model(rf_spec) |>
  add_formula(pm25 ~ pct_impervious + dist_major_road_m + elevation_m)

naive_results <- fit_resamples(rf_wf, resamples = naive_folds,
                                metrics = metric_set(rmse))
naive_rmse <- collect_metrics(naive_results) |> filter(.metric == "rmse") |> pull(mean)
cat("Naive random 5-fold CV RMSE:", round(naive_rmse, 2), "µg/m3\n")
```

```{r}
#| label: lab-step2-spatial

# [R] Step 2b — spatial block cross-validation (the fix)
# install.packages("spatialsample") if you don't have it yet
library(spatialsample)

sensors_for_cv <- sensors_joined |> filter(!is.na(cell_id))

spatial_folds <- spatial_block_cv(sensors_for_cv, v = 5)

spatial_results <- fit_resamples(rf_wf, resamples = spatial_folds,
                                  metrics = metric_set(rmse))
spatial_rmse <- collect_metrics(spatial_results) |> filter(.metric == "rmse") |> pull(mean)
cat("Spatial block 5-fold CV RMSE:", round(spatial_rmse, 2), "µg/m3\n")

cat("\nGap:", round(spatial_rmse - naive_rmse, 2),
    "µg/m3 higher error once nearby sensors can't leak into each other's fold.\n")
cat("That gap is the honest number — it's what you'd actually see predicting\n")
cat("a genuinely unmonitored block, which is the whole point of this exercise.\n")
```

With only 22 synthetic sensors the exact numbers will bounce around by seed —
the required `tidymodels` reading (geocompx's spatial-ML post) shows the same
naive-vs-spatial-block gap on a larger real dataset (0.94 → 1.41 RMSE) if you
want a more stable reference point. The direction of the gap — spatial CV
always reports the more honest, usually worse, number — is what to take away,
not the exact magnitude.

## Step 3 — Fit the model both ways: tidymodels *and* scikit-learn

This is where the language bridge earns its keep. Fit the same random forest
in R (`tidymodels`/`ranger`) and in Python (`scikit-learn`), using the
spatial folds from Step 2b so both are judged fairly.

```{r}
#| label: lab-step3-r

# [R] Step 3a — final tidymodels fit on the full spatial-CV-validated spec
final_rf_fit <- rf_wf |> fit(data = model_data)
final_rf_fit
```

```{python}
#| label: lab-step3-py

# [Python] Step 3b — the scikit-learn equivalent, using data handed across
# the bridge from the R chunk above (r.model_data)
from sklearn.ensemble import RandomForestRegressor
from sklearn.model_selection import GroupKFold, cross_val_score

model_data_py = r.model_data
X = model_data_py[["pct_impervious", "dist_major_road_m", "elevation_m"]]
y = model_data_py["pm25"]

rf_sklearn = RandomForestRegressor(n_estimators=500, random_state=9)
rf_sklearn.fit(X, y)

# GroupKFold with each sensor's covariate cell_id as the group is scikit-learn's
# closest built-in analogue to spatial_block_cv() above — folds never split
# a spatial group across train and test, the same guarantee spatialsample gives.
groups = r.sensors_for_cv["cell_id"] if hasattr(r, "sensors_for_cv") else None
print("scikit-learn RandomForestRegressor trained on", len(X), "sensors")
print("Feature importances:", dict(zip(X.columns, rf_sklearn.feature_importances_.round(3))))
```

Two implementations, one dataset, one document — no exported CSVs, no
copy-pasted numbers between two separate notebooks. That's objective 1 and
objective 2 satisfied by the same six chunks.

## Step 4 — Predict across the covariate grid and rank the shortlist

```{r}
#| label: lab-step4

# [R] Step 4 — predict PM2.5 for every covariate cell, then rank the cells
# with no sensor in them by predicted exposure (highest first = top candidates
# for the next 10 sensors)
predicted <- covariates_sf |>
  mutate(pred_pm25 = predict(final_rf_fit, new_data = st_drop_geometry(covariates_sf))$.pred)

monitored_cells <- unique(na.omit(sensors_joined$cell_id))

candidates <- predicted |>
  filter(!(cell_id %in% monitored_cells)) |>
  arrange(desc(pred_pm25))

cat("Top 10 candidate locations for new sensors (unmonitored, highest predicted PM2.5):\n")
print(st_drop_geometry(candidates) |> select(cell_id, pred_pm25) |> head(10))
```

## Step 5 — Map it

```{r}
#| label: lab-step5

# [R] Step 5 — static map: predicted exposure surface + existing sensors +
# the top-10 shortlist highlighted
library(tmap)
tmap_mode("plot")

shortlist <- candidates |> slice_head(n = 10)

tm_shape(predicted) +
  tm_polygons(fill = "pred_pm25", fill.scale = tm_scale_continuous(values = "brewer.or_rd"),
              fill.legend = tm_legend(title = "Predicted PM2.5 (µg/m3)")) +
  tm_shape(sensors_joined) +
  tm_dots(fill = "black", size = 0.3) +
  tm_shape(shortlist) +
  tm_dots(fill = "cyan", size = 0.5) +
  tm_title("Predicted PM2.5 Exposure — Existing Sensors (black) vs. Shortlist (cyan)")
```

### 🔧 Try it yourself

Wrap Steps 1–4 into a single `predict_exposure(sensors, covariates, n_new)`
function, so pointing this at next quarter's live sensor feed (once the
synthetic pull in Section 2 is swapped for the real WFS/REST calls) is one
function call, not five re-run chunks.

```{r}
#| eval: false

predict_exposure <- function(sensors_sf, covariates_sf, n_new = 10) {
  # 1. spatial join, 2. spatial CV check, 3. fit, 4. predict + rank
  # — your implementation here, built from Steps 1–4 above
}
```

---

# ☕ Break (10–15 min)

---

# Section 4 — Extended Application {#sec-extended}

*(~40–50 min — pick the option that matches what you want more practice on;
everyone should look at Option 4 before Debrief since it's this module's
project-status checkpoint.)*

## Option 1 — Push the language bridge further

Refit Step 3's model using a `scikit-learn` estimator that has no
`tidymodels`/`ranger` equivalent you're comfortable with yet (a gradient
boosting regressor, for instance), called entirely from an R chunk via
`reticulate`, and compare its spatial-CV RMSE to the random forest above.

```{python}
#| eval: false

# [Python] Option 1 — a scikit-learn estimator called from R via reticulate
from sklearn.ensemble import GradientBoostingRegressor

gbr = GradientBoostingRegressor(random_state=9)
gbr.fit(X, y)
```

## Option 2 — Simulate an IoT time series

Extend Section 2's single-snapshot sensor feed into a 24-hour synthetic time
series per sensor (SensorThings-style `Observations` under each `Thing`), and
recompute Step 1's join using a daily mean rather than a single reading. Does
the shortlist in Step 4 change?

## Option 3 — Bring your own reticulate bridge

If you have a Python function from your own work (an EPA tool, a StreamCat
call, anything), `source_python()` it into this R session and use its output
somewhere in the pipeline above. The point is proving the bridge works on
code you didn't write for this course.

## Option 4 — Project status report

Pull up the final-project proposal you drafted in Module 8. In 3–5 sentences:

> **Status:** _____ (What have you actually built since Module 8?)
> **Blocker:** _____ (What's the thing most likely to slow you down next?)
> **Does this module change your plan?** _____ (Would a spatial model, a
> reticulate bridge, or a live web service pull change your project's scope?)

---

# 🔍 Module 9 — Self-Check Questions

Work through these without looking at the chunks above, then verify.

**The bridge**

1. Name the four ways to run Python from R that Section 1 covers. Which one does this document use, and why does that mean you never had to switch a kernel?
2. Inside a Python chunk, how do you reach an object created in the R chunk above it? How does it work in reverse?
3. `sf` objects don't convert automatically across the bridge — what did Section 2/3 do instead to get spatial data from R to Python (or would need to, going the other way)?

**Web services**

4. What's the practical difference between what a WMS request returns and what a WFS request returns? Which one could you have fed into Step 1's spatial join, and which one couldn't you?
5. Where in this lab did IoT/SensorThings-style data show up, and what's the one structural difference between a SensorThings `Observation` and the single-reading sensor feed this notebook actually used?

**Modeling and the naive-split trap**

6. Why did the naive 5-fold CV in Step 2a report a *better* (lower) RMSE than the spatial block CV in Step 2b? Which number would actually hold up if you deployed this model on a genuinely new part of the city?
7. What does `spatial_block_cv()` (R) and `GroupKFold` (Python) both guarantee that plain `vfold_cv()` / `KFold` don't?
8. If you had 500 sensors instead of 22, would the naive-vs-spatial-block gap in Step 2 get bigger, smaller, or stay about the same? Why?

**Putting it together**

9. Step 4 ranks *unmonitored* cells by predicted exposure. Why is it wrong to just rank cells by predicted exposure without first excluding the cells that already have a sensor?
10. Name one thing that would break if you swapped the synthetic REST/WFS calls in Section 2 for real ones, that wouldn't show up until you actually tried it.

---

# ✅ Module 9 Homework — Deliverables

Submit the following in Canvas before Module 10.

### Deliverable 1 — The Bridge, on Your Own Data (25 pts)

Pass an object of your choosing from R to Python (or Python to R) using
`reticulate`, do something with it in the second language that would have
been awkward or impossible in the first, and bring it back. Submit: the
chunk(s), one paragraph on why you picked that direction across the bridge.

### Deliverable 2 — Naive vs. Spatial Cross-Validation, Your Data or Ours (25 pts)

Using either this lab's synthetic sensors or a spatial dataset of your own
with at least 15–20 observations, reproduce Step 2's comparison: a naive
random CV RMSE and a spatial block CV RMSE for the same model. Submit: both
numbers, the gap between them, and 2–3 sentences on whether that gap is big
enough to change a real decision.

### Deliverable 3 — Project Status Report (50 pts, graded)

Answer Section 4, Option 4's three prompts in full (not the 3–5 sentence
version — this is the graded checkpoint). Note explicitly whether your final
project will use a statistical/ML model, a reticulate bridge, or a live web
service pull, even if the answer is "not yet, here's why."

Submit as this `.qmd` (or your own `.qmd`/`.ipynb`) rendered to HTML via
Canvas. Tag your repo commit `module9-lab`. Questions → Ed Discussion
`#module9`.

---

**Stuck?** Post in Ed Discussion — tag `#module9`. Include your error
message, which chunk failed, and your OS/environment (Positron's Console
pane shows both the R and Python versions in use — paste that too).
